DNAm Hierarchical Model - Prior Sensitivity Analysis 

### 2020-10-02 conserved_genes_and_paths.Rmd
updated prior sensitivity analysis

knit_exit()

@@

  # convert to CDF, compute K-S-like statistic
  v <- cumsum(runsum)
  score <- max(c(0,v)) + min(c(0,v))
  GSEA_df$score[iset] <- score
}
cat(proc.time()-ptm)
cat(sprintf("Processed %i genes, %i pathways/genesets. Begin bootstrapping...",ngene,nsets))

@@

shuffle_genesets <- function(n_geneset, n_rep, data) {
  bootstrap = numeric(n_rep)
  for (ishuf in 1:n_rep) {
    # for random shuffles just make a random logical mask
    inds = sample(1:ngene, n_geneset, replace = F)
    inmask = logical(ngene)
    inmask[inds] = TRUE
    runsum = evidence/sum(evidence[inmask]) * inmask - 1/(ngene - length(set)) * !inmask

    # convert to CDF, compute K-S-like statistic
    v <- cumsum(runsum)
    score <- max(c(0, v)) + min(c(0, v))
    bootstrap[ishuf] <- score
  }
  bootstrap
}

names(setlengths) <- setlengths
cat(sprintf("Shuffling %i set-lengths: 1000x random shuffles (w/o replacement)",length(setlengths)))
ptm<-proc.time()
res <- lapply(setlengths, shuffle_genesets, n_rep = 1000, data = evidence)
proc.time()-ptm

>>>>

@@

  # convert to CDF, compute K-S-like statistic
  v <- cumsum(runsum)
  #score <- max(v) # GSEA score
  score <- max(c(0,v)) + min(c(0,v)) # GSVA score
  GSEA_df$score[iset] <- score
}
cat(proc.time()-ptm)
cat(sprintf("Processed %i genes, %i pathways/genesets. Begin bootstrapping...",ngene,nsets))

@@

shuffle_genesets <- function(n_geneset, n_rep, data) {
  bootstrap = numeric(n_rep)
  for (ishuf in 1:n_rep) {
    # for random shuffles just make a random logical mask
    inds = sample(1:ngene, n_geneset, replace = T)
    inmask = logical(ngene)
    inmask[inds] = TRUE
    runsum = evidence/sum(evidence[inmask]) * inmask - 1/(ngene - length(set)) * !inmask

    # convert to CDF, compute K-S-like statistic
    v <- cumsum(runsum)
    #score <- max(v) # GSEA score
    score <- max(c(0, v)) + min(c(0, v)) # GSVA score
    bootstrap[ishuf] <- score
  }
  bootstrap
}

names(setlengths) <- setlengths
cat(sprintf("Shuffling %i set-lengths: 5000x random shuffles (w/ replacement)",length(setlengths)))
ptm<-proc.time()
res <- lapply(setlengths, shuffle_genesets, n_rep = 5000, data = evidence)
proc.time()-ptm

### 2020-10-02 prior_sensitivity.Rmd
updated prior sensitivity analysis

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
library(gtools)

@@

select <- dplyr::select
rename <- dplyr::rename
filter <- dplyr::filter
slice <- dplyr::slice

set.seed(1234)
```

```{r prior_sensitivity, echo=FALSE}

for (i in c(1:4)) {
  stanFile = "FullResultsTCGA_relaxgamma3.Rdata"
  stanFile = switch(i,
                    "FullResultsTCGA_gamma.Rdata",
                    "FullResultsTCGA_relaxgamma2.Rdata",
                    "FullResultsTCGA_relaxgamma3.Rdata",
                    "FullResultsTCGA_relaxgamma5.Rdata")
  print(stanFile)
  #load(here(stanFile))
}

>>>>

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(tidyverse)
library(knitr)
#library(gtools)
library(emdist)

@@

select <- dplyr::select
rename <- dplyr::rename
filter <- dplyr::filter
slice <- dplyr::slice
```

```{r prior_sensitivity, echo=FALSE, comment=NA}

n_eff_mat <- data.frame(Params=c("mu","betaT","sigmaP","sigmaPT","sigmaT","sigmaE","lp"))
Rhat_mat <- n_eff_mat
KLdiv_mat <- n_eff_mat %>% slice(1:6)
BCdist_mat <- KLdiv_mat
EMdist_mat <- KLdiv_mat

for (i in c(1:4)) {
  stanFile = "FullResultsTCGA_relaxgamma3.Rdata"
  stanFile = switch(i,
                    "FullResultsTCGA_gamma.Rdata",
                    "FullResultsTCGA_relaxgamma2.Rdata",
                    "FullResultsTCGA_relaxgamma3.Rdata",
                    "FullResultsTCGA_relaxgamma5.Rdata")
  print(stanFile)
  load(here(stanFile))
  print("loaded")

  n_eff_temp <-
  data.frame(
  mu = mu_full$n_eff,
  betaT = betaT_full$n_eff,
  sigmaP = sigmaP_full$n_eff,
  sigmaPT = sigmaPT_full$n_eff,
  sigmaT = sigmaT_full$n_eff,
  sigmaE = sigmaE_full$n_eff,
  lp = lp_full$n_eff
  ) %>% gather(param,value,mu:lp) %>% group_by(param) %>% summarize(
  mean = mean(value),
  .groups = "drop"
  ) %>% arrange(match(.$param,c("mu","betaT","sigmaP","sigmaPT","sigmaT","sigmaE","lp")))

  Rhat_temp <-
  data.frame(
  mu = mu_full$Rhat,
  betaT = betaT_full$Rhat,
  sigmaP = sigmaP_full$Rhat,
  sigmaPT = sigmaPT_full$Rhat,
  sigmaT = sigmaT_full$Rhat,
  sigmaE = sigmaE_full$Rhat,
  lp = lp_full$Rhat
  ) %>% gather(param,value,mu:lp) %>% group_by(param) %>% summarize(
  mean = mean(value),
  .groups = "drop"
  ) %>% arrange(match(.$param,c("mu","betaT","sigmaP","sigmaPT","sigmaT","sigmaE","lp")))

  # KL divergence
  # formula: KL(P||Q) = sum(P(i) * log( P(i)/Q(i) )), where P(i),Q(i) = density
  KL_temp <- numeric(6)
  #Bhattacharya distance
  # formula: BC = sum( sqrt(P(i)*Q(i)) )
  BC_temp <- numeric(6)
  # Earth-movers (Wasserstein) distance
  EM_temp <- numeric(6)

  for (j in c(1:6)) {
    # choose parameter
    data <- switch(j,
                   mu_full$p50,
                   betaT_full$p50,
                   sigmaP_full$p50,
                   sigmaPT_full$p50,
                   sigmaT_full$p50,
                   sigmaE_full$p50)
    breaks <- switch(j,
                     seq(-5,5,0.1),
                     seq(-6,6,0.12), #5.5,.11
                     seq(0,3,0.03), #2.6
                     seq(0,3,0.03),
                     seq(0,3,0.03), #1.2
                     seq(0,3,0.03)) #2.5
    P <- hist(data, plot=FALSE, breaks=101) # run hist to get density and midpoints
    x <- P$mids
    # define prior density by pre-defined priors at P midpoints
    if (i==1){ # non relaxed (1x)
      priordens <- switch(j,
                      0.3557*dnorm(x, -3.0675, 0.75214) + 0.6443*dnorm(x, 1.1290, 1.3920),
                      dcauchy(x, -0.04051, 0.19330),
                      dgamma(x, 5.1989, 16.6192),
                      dgamma(x, 3.5539, 5.5324),
                      dgamma(x, 6.8383, 25.7560),
                      dgamma(x,2,2))
    } else if (i==2) { # 2x relaxed
      priordens <- switch(j,
                      0.3557*dnorm(x, -3.0675, 1.0637) + 0.6443*dnorm(x, 1.1290, 1.9685),
                      dcauchy(x,-0.04051, 0.38660),
                      dgamma(x, 3.4016, 9.5057),
                      dgamma(x, 2.5210, 3.2948),
                      dgamma(x, 4.2574, 14.370),
                      dgamma(x,2,2))
    } else if (i==3) { # 3x relaxed
      priordens <- switch(j,
                      0.3557*dnorm(x, -3.0675, 1.3027) + 0.6443*dnorm(x, 1.1290, 2.4110),
                      dcauchy(x,-0.04051, 0.5799),
                      dgamma(x, 2.7693, 7.0029),
                      dgamma(x, 2.1457, 2.4819),
                      dgamma(x, 3.3643, 10.430),
                      dgamma(x,2,2))
    } else if (i==4) { # 5x relaxed
      priordens <- switch(j,
                      0.3557*dnorm(x, -3.0675, 1.6818) + 0.6443*dnorm(x, 1.1290, 3.1125),
                      dcauchy(x,-0.04051, 0.96650),
                      dgamma(x, 2.2298, 4.8674),
                      dgamma(x, 1.8166, 1.7689),
                      dgamma(x, 2.6144, 7.1221),
                      dgamma(x,2,2))
    }
    Pi <- P$density[which(P$density>0)]
    Qi <- priordens[which(P$density>0)]
    KL_temp[j] = sum(Pi * log(Pi/Qi))

    # BCD
    BC_temp[j] = -log(sum(sqrt(P$density*priordens)))

    # EMD
    A = cbind(P$density,P$mids)
    B = cbind(priordens,P$mids)
    EM_temp[j] = emd(A,B)

    # code below is for plotting data vs priors
    #Q <- P
    #Q$density <- priordens
    #Q$counts <- Q$density*866091*(Q$breaks[2]-Q$breaks[1])
    #c1 <- rgb(173,216,230,max = 255, alpha = 80, names = "lt.blue")
    #c2 <- rgb(255,192,203, max = 255, alpha = 80, names = "lt.pink")
    #plot(P,col=c1, main=paste("i=",i,", j=",j))
    #plot(Q,col=c2,add=TRUE)
  }

  varname = switch(i, "1x", "2x", "3x", "5x")
  n_eff_mat <- n_eff_mat %>% mutate(!!varname := n_eff_temp$mean)
  Rhat_mat <- Rhat_mat %>% mutate(!!varname := Rhat_temp$mean)
  KLdiv_mat <- KLdiv_mat %>% mutate(!!varname := KL_temp)
  BCdist_mat <- BCdist_mat %>% mutate(!!varname := BC_temp)
  EMdist_mat <- EMdist_mat %>% mutate(!!varname := EM_temp)

}

print("N_eff")
kable(n_eff_mat, digits=2)
print("Rhat")
kable(Rhat_mat, digits=3)
print("KLdiv")
kable(KLdiv_mat, digits=3)
cat("\newpage")
print("BCdist")
kable(BCdist_mat, digits=3)
print("EMdist")
kable(EMdist_mat, digits=3)
